### 2024-05-22 getting-started.qmd
20240522 - clean up formatting
20240522 - update computations

>>>>

For instance, you can use the following code to install the `nflreadr` package:

install.packages("nflreadr")
```

## Load Packages {#sec-loadPackages}

```{r}
library("nflreadr")
library("nflfastR")
library("progressr")
library("tidyverse")
```

## Download Football Data {#sec-downloadFootballData}

### Player Info {#sec-downloadPlayerInfo}

### Play-By-Play Data {#sec-downloadPlayByPlay}

To download play-by-play data from prior weeks and seasons, we can use the `load_pbp()` function of the `nflreadr` package.
We add a progress bar using the `with_progress()` function from the `progressr` package because it takes a while to run.

::: {#nte-downloadPlayByPlay .callout-note title="Downloading play-by-play data"}
Note: the following code takes a while to run.
:::

nfl_pbp <- progressr::with_progress(
  nflreadr::load_pbp(seasons = TRUE))
```

```{r}
#| include: false

nfl_pbp <- nflreadr::load_pbp(seasons = TRUE)
```

```{r}
#| eval: false
#| include: false

library("here")

save(
  nfl_pbp,
  file = here("data", "nfl_pbp.RData")
)
```

```{r}
#| include: false

load(file = here("data", "nfl_pbp.RData"))
```

### Historical Weekly Actual Player Statistics {#sec-downloadActualStats}

We can download historical week-by-week actual player statistics using the `load_player_stats()` function from the `nflreadr` package:

nfl_actualStats_offense_weekly <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    stat_type = "offense"))

nfl_actualStats_defense_weekly <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    stat_type = "defense"))

nfl_actualStats_kicking_weekly <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    stat_type = "kicking"))
```

```{r}
#| eval: false
#| include: false

save(
  nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
  file = here("data", "nfl_actualStats_weekly.RData")
)
```

```{r}
#| include: false

load(file = here("data", "nfl_actualStats_weekly.RData"))
```

## Calculations {#sec-calculations}

### Historical Actual Player Statistics {#sec-calculateActualStats}

In addition to week-by-week actual player statistics, we can also compute historical actual player statistics as a function of different timeframes, including season-by-season and career statistics.

#### Career Statistics {#sec-calculateActualCareerStats}

First, we can compute the players' career statistics using the `calculate_player_stats()`, `calculate_player_stats_def()`, and `calculate_player_stats_kicking()` functions from the `nflfastR` package for offensive players, defensive players, and kickers, respectively.

::: {#nte-calculateActualCareerStats .callout-note title="Calculating players' career statistics"}
Note: the following code takes a while to run.
:::

nfl_actualStats_offense_career <- nflfastR::calculate_player_stats(
  nfl_pbp,
  weekly = FALSE)

nfl_actualStats_defense_career <- nflfastR::calculate_player_stats_def(
  nfl_pbp,
  weekly = FALSE)

nfl_actualStats_kicking_career <- nflfastR::calculate_player_stats_kicking(
  nfl_pbp,
  weekly = FALSE)
```

```{r}
#| eval: false
#| include: false

save(
  nfl_actualStats_offense_career, nfl_actualStats_defense_career, nfl_actualStats_kicking_career,
  file = here("data", "nfl_actualStats_career.RData")
)
```

```{r}
#| include: false

load(file = here("data", "nfl_actualStats_career.RData"))
```

#### Season-by-Season Statistics {#sec-calculateActualSeasonalStats}

Second, we can compute the players' season-by-season statistics.

```{r}
seasons <- unique(nfl_pbp$season)

nfl_pbp_seasonalList <- list()
nfl_actualStats_offense_seasonalList <- list()
nfl_actualStats_defense_seasonalList <- list()
nfl_actualStats_kicking_seasonalList <- list()
```

::: {#nte-calculateActualSeasonalStats .callout-note title="Calculating players' season-by-season statistics"}
Note: the following code takes a while to run.
:::

pb <- txtProgressBar(
  min = 0,
  max = length(seasons),
  style = 3)

for(i in 1:length(seasons)){
  # Subset play-by-play data by season
  nfl_pbp_seasonalList[[i]] <- nfl_pbp %>% 
    filter(season == seasons[i])

  # Compute actual statistics by season
  nfl_actualStats_offense_seasonalList[[i]] <- 
    nflfastR::calculate_player_stats(
      nfl_pbp_seasonalList[[i]],
      weekly = FALSE)

  nfl_actualStats_defense_seasonalList[[i]] <- 
    nflfastR::calculate_player_stats_def(
      nfl_pbp_seasonalList[[i]],
      weekly = FALSE)

  nfl_actualStats_kicking_seasonalList[[i]] <- 
    nflfastR::calculate_player_stats_kicking(
      nfl_pbp_seasonalList[[i]],
      weekly = FALSE)

  nfl_actualStats_offense_seasonalList[[i]]$season <- seasons[i]
  nfl_actualStats_defense_seasonalList[[i]]$season <- seasons[i]
  nfl_actualStats_kicking_seasonalList[[i]]$season <- seasons[i]

  print(
    paste("Completed computing projections for season: ", seasons[i], sep = ""))

  # Update the progress bar
  setTxtProgressBar(pb, i)
}

# Close the progress bar
close(pb)

nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonalList %>% 
  bind_rows()
nfl_actualStats_defense_seasonal <- nfl_actualStats_defense_seasonalList %>% 
  bind_rows()
nfl_actualStats_kicking_seasonal <- nfl_actualStats_kicking_seasonalList %>% 
  bind_rows()
```

```{r}
#| eval: false
#| include: false

save(
  nfl_actualStats_offense_seasonal, nfl_actualStats_defense_seasonal, nfl_actualStats_kicking_seasonal,
  file = here("data", "nfl_actualStats_seasonal.RData")
)
```

```{r}
#| include: false

load(file = here("data", "nfl_actualStats_seasonal.RData"))
```

#### Week-by-Week Statistics {#sec-calculateActualWeeklyStats}

We already load players' week-by-week statistics [above](#downloadActualStats).
Nevertheless, we could compute players' weekly statistics from the play-by-play data using the following syntax:

nfl_actualStats_weekly <- nflfastR::calculate_player_stats(
  nfl_pbp,
  weekly = TRUE)
```

### Historical Actual Fantasy Points {#sec-calculateActualPoints}
